---
title: "Initial Korean covid 19"
author: "Jennifer Epinoza "
date: "Summer 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this 

```{r}
kr3 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
```


  
  
```{r}
basicCC <- ggplot(kr3,aes(x=sex,y=count,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

```{r}
basicCC <- ggplot(kr3,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

When you look at the chart percentage you can see that the males are dying more. about 65% of the dead column is made up of males. 

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

When looking at the numerical results we can see that 64 out of 100 males are dead and and 36 out of 100 females are dead. with this we can see that there are more fales dead than females. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

based off of this we can see that covid is killing more males than females and this was seen statistically from the results.

## Conclusions

After testing all of the tests we came to the conclusion that in percentage wise there was more females in the alive section than males. also percentage wise males dominated the dead section.This showed that the death toll was not equivalent between males and females. Therefore there was more death seen in males than females.
